% !TeX root = ../main.tex
% Add the above to each chapter to make compiling the PDF easier in some editors.

\chapter{Introduction}\label{chapter:introduction}

Tensor network contractions have become a core computational primitive in fields ranging from quantum many-body physics and quantum circuit simulation to chemistry and high-dimensional machine learning and data analysis. Their appeal comes from the ability to represent and manipulate exponentially large objects through structured factorizations, often making previously intractable problems accessible in practice \cite{evenbly2014improving}. Yet this promise comes with a major obstacle: the cost of contracting a tensor network can grow dramatically depending on how intermediate tensors are formed. In many realistic settings, the limiting factor is not only the number of arithmetic operations, but the peak memory required to store these intermediates. A contraction strategy that is theoretically efficient in \ac{FLOPS} may still be unusable in practice if it creates tensors too large to fit in memory.

This makes memory reduction a central problem rather than a secondary implementation concern. Prior work on contraction ordering has already shown that the sequence in which tensors are contracted has a decisive impact on computational feasibility and overall performance \cite{pfeifer2014faster,schindler2020algorithms}. At the same time, high-performance tensor contraction frameworks and tensor-transposition libraries have demonstrated that practical efficiency depends heavily on how contractions are mapped onto hardware, how data is rearranged, and how temporary storage is managed \cite{hirata2003tensor,DBLP:journals/corr/SpringerB16}. In other words, optimizing tensor networks requires attention not only to algebraic structure, but also to memory footprint, data movement, and layout-aware execution.

The importance of optimizing tensor network contractions originates from both theoretical and practical considerations. From a theoretical perspective, tensor contraction is known to be computationally hard in general, with complexity that can grow exponentially with the size of the network \cite{pfeifer2014faster}. From a practical standpoint, modern applications such as quantum circuit simulation and many-body physics rely heavily on large-scale tensor contractions, where inefficiencies quickly become restrictive.

A key observation in \ac{HPC} is that memory bandwidth and data movement increasingly dominate execution time, often surpassing the cost of \ac{FLOPS} \cite{hirata2003tensor,huang2021efficient}. In large-scale simulations, such as those targeting quantum supremacy benchmarks \cite{springer2017hptt,huang2021efficient}, the ability to efficiently manage memory and minimize data movement is critical to achieving feasible runtimes.

Moreover, the peak memory usage during tensor contraction directly determines whether a computation is even possible on a given system. Many tensor network contractions fail not due to computational limits, but because intermediate tensors exceed available memory. As a result, reducing peak memory usage is not merely an optimization objective, but a fundamental requirement for scalability.

Despite these challenges, most existing optimization techniques primarily focus on reducing computational cost (\ac{FLOPS}) through contraction ordering, often overlooking memory-related factors such as tensor layout, data locality, and permutation overhead. This creates a gap between theoretical optimization and practical performance, particularly on modern architectures where memory hierarchy plays a central role.

\section{Problem Formulation and Challenges}

Ultimately, the goal of contraction is to compute a final tensor by successively contracting pairs of tensors along shared indices. This process involves three key components: the contraction order, which determines the sequence in which tensor pairs are combined; the intermediate tensor shapes, which define the sizes and dimensionalities of tensors generated during the contraction; and the tensor layout and permutation, which determine how tensor indices are arranged in memory and therefore how efficiently operations can be executed.

While contraction ordering has been extensively studied, the challenges associated with intermediate tensor shapes and tensor layouts remain significant and often underexplored. In particular, tensor permutations play a crucial role in enabling efficient computation. High-performance implementations typically rely on transforming tensor contractions into matrix multiplications via reshaping and permutation, a technique known as Tensor-Tensor Generalized Multiplication (TTGT) \cite{DBLP:journals/corr/SpringerB16}.

However, these permutations introduce additional memory movement, which can be costly. In many cases, tensors are repeatedly permuted and reshaped throughout the contraction sequence. This leads to increased memory bandwidth usage, poor cache locality, and higher peak memory consumption due to the creation of intermediate copies. These effects collectively degrade performance and can become a dominant bottleneck in large-scale computations.

Furthermore, existing approaches typically treat permutations as local operations required to enable individual contractions. This local perspective ignores the global structure of the contraction sequence and can result in redundant or suboptimal data transformations that accumulate over the course of execution \cite{brandejs2026tensor}.

\section{Research Aim}

This thesis proposes a complementary approach that shifts part of the optimization focus to the \textit{post-contraction phase}. Instead of solely optimizing the contraction order, we investigate how global tensor permutations can be applied after determining a contraction path to improve memory efficiency.

The key idea is to treat tensor layout as a first-class optimization variable. By analyzing the entire contraction sequence, it becomes possible to reduce redundant permutations across consecutive contractions, improve data locality and cache utilization, and minimize peak memory usage by carefully controlling intermediate tensor layouts.

This approach is relatively niche compared to traditional methods, as most existing work focuses on pre-contraction optimization and local permutation strategies. However, given the growing importance of memory efficiency in modern computing systems, global permutation optimization presents a promising direction for further research.

\section{Contributions of This Work}

In summary, this thesis makes several key contributions. It introduces a novel framework for post-contraction optimization based on global tensor permutations. It provides a detailed analysis of the impact of tensor layouts on memory usage and data movement. It highlights the limitations of existing pre-contraction optimization techniques in addressing memory-related challenges. Finally, it demonstrates how tensor permutation strategies can complement contraction ordering to achieve improved overall performance.

By bridging the gap between contraction planning and memory-aware tensor layout optimization, this work aims to provide a more comprehensive approach to efficient tensor network contraction.

