\addcontentsline{toc}{chapter}{Acknowledgments}
\thispagestyle{empty}

\vspace*{20mm}

\begin{center}
    {\usekomafont{sectioning}\usekomafont{section} Acknowledgments}
\end{center}

\vspace{10mm}

Completing this thesis has been one of the most rewarding experiences of my academic journey. It would not have been possible without the support, guidance, and encouragement of many remarkable people.

First and foremost, I would like to extend my sincere thanks to my supervisor, \textbf{Manuel Geiger}, whose expertise, mentorship, and countless discussions were instrumental in shaping this thesis.

I would, also, like to express my deepest gratitude to my examiner, \textbf{Dr. Mohammed Shahzad}, for his invaluable guidance, insightful feedback, and continuous encouragement throughout this research. 

To my friends, thank you for being a constant source of motivation, laughter, and perspective. Whether through late-night conversations, words of encouragement, travels, or simply being there, your support made this journey far less daunting.

To my family, words cannot fully express my gratitude. Thank you for believing in me even during the moments when I doubted myself. Your unconditional love, sacrifices, patience, and unwavering support have been the foundation upon which every one of my achievements has been built. Everything I have accomplished is a reflection of the opportunities and values you have given me.

Finally, to that special person who stood beside me through every high and every low, thank you. Your unwavering support, endless patience, and constant belief in me carried me through this journey. You celebrated every success, comforted every setback, and reminded me to keep moving forward when the path seemed uncertain. Having someone who remained by my side regardless of the circumstances is a gift I will always cherish, and I am profoundly grateful for your presence in my life.

To everyone who has been part of this journey or crossed my path, thank you. This achievement belongs not only to me, but also to you :)
\cleardoublepage{}
